---
title: "ESM204_HW4"
author: "Annie Combs, Alissa Patterson, Michelle Geldin"
date: "May 16, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
```

```{r}
warming <- read_csv("warming.csv") %>%
  clean_names()


damages <- read_csv("damages.csv")%>%
  clean_names()

model_damages <- lm(damages  ~ warming, data=damages)

summary(model_damages)

```
## 1. Using damages.csv, estimate a quadratic damage function relating the dollar value of damages to the change in global mean temperature. Estimate a model with a zero intercept because damages by construction must equal zero when there is no climate change. Plot your estimated damage function, overlaid with a scatterplot of the underlying data.

```{r}
quadratic <- function(x) {y = x^2
return(y)}

warming2 <- warming^2

damages_lm <- lm(damages = warming + warming2, data = damages)

damage_function <- quadratic(damages)

ggplot(damages_lm, aes(x = warming, y = damages))+
  geom_line()

```

## 2. Use warming.csv and your estimated damage function to predict damages in each year under the baseline climate and the pulse scenario. Make four plots: (1) damages over time without the pulse, (2) damages over time with the pulse, (3) the difference in damages over time that arises from the pulse, and (4) the difference in damages over time from the pulse per ton of
 
```{r}
# storing a function for damages from warming based on the quadratic model
damages <- function(warming) {
  damages <- damages_lm$coefficients[1]*warming + damages_lm$coefficients[2]*(warming^2)
  return(damages)
}

pulse <- warming %>% 
  mutate(damages_baseline = damages(warming_baseline)) %>% 
  mutate(damages_pulse = damages(warming_pulse)) %>% 
  mutate(difference = damages_pulse - damages_baseline) %>% 
  mutate(damage_per_ton = difference/35e+9)

p1 <- ggplot(data = pulse, aes(x = year, y = damages_baseline/1e+12)) +
  geom_line(size = 1.2, color = "firebrick2") +
  labs(x = "Year", y = "Damages (trillions USD)", title = "Baseline damages") +
  theme_minimal(14) +
  theme(plot.title = element_text(hjust = 0.48, face = "bold"))

p1

p2 <- ggplot(data = pulse, aes(x = year, y = damages_pulse/1e+12)) +
  geom_line(size = 1.2, color = "firebrick2") +
  labs(x = "Year", y = "Damages (trillions USD)", title = "Damages with pulse") +
  theme_minimal(14) +
  theme(plot.title = element_text(hjust = 0.48, face = "bold"))

p2

p3 <- ggplot(data = pulse, aes(x = year, y = difference/1e+09)) +
  geom_line(size = 1.2, color = "firebrick2") +
  labs(x = "Year", y = "Damages (billions USD)", title = "Additional damages from pulse") +
  theme_minimal(14) +
  theme(plot.title = element_text(hjust = 0.48, face = "bold"))

p3

p4 <- ggplot(data = pulse, aes(x = year, y = difference/35e+9)) +
  geom_line(size = 1.2, color = "firebrick2") +
  labs(x = "Year", y = "Damages per ton CO2 (USD)", title = "Per ton of CO2") +
  theme_minimal(14) +
  theme(plot.title = element_text(hjust = 0.48, face = "bold"))

p4

```
 
 
## 3. The SCC is the present discounted value of the stream of future damages caused by one additional ton of CO2. The Obama Administration used a discount rate of 3% to discount damages. Recently, New York State used a discount rate of 2%. Calculate and make a plot of the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates. Explain the intuition for how the discount rate affects the SCC.

```{r}
scc <- pulse %>% 
  mutate(years_into_future = year-2022) %>% 
  mutate(scc_not_discounted = difference/35e+9) %>% 
  mutate(dr0.25 = scc_not_discounted/(1.0025^years_into_future)) %>% 
  mutate(dr0.5 = scc_not_discounted/(1.005^years_into_future)) %>% 
  mutate(dr0.75 = scc_not_discounted/(1.0075^years_into_future)) %>% 
  mutate(dr1 = scc_not_discounted/(1.01^years_into_future)) %>% 
  mutate(dr1.5 = scc_not_discounted/(1.015^years_into_future)) %>% 
  mutate(dr2 = scc_not_discounted/(1.02^years_into_future)) %>% 
  mutate(dr3 = scc_not_discounted/(1.03^years_into_future)) %>% 
  mutate(dr4 = scc_not_discounted/(1.04^years_into_future)) %>% 
  mutate(dr5 = scc_not_discounted/(1.05^years_into_future)) %>% 
  mutate(dr6 = scc_not_discounted/(1.06^years_into_future)) %>% 
  mutate(dr2.1 = scc_not_discounted/(1.021^years_into_future))
  

scc0.025 <- sum(scc$dr0.25)
scc0.5 <- sum(scc$dr0.5)
scc0.75 <- sum(scc$dr0.75)
scc1 <- sum(scc$dr1)
scc1.5 <- sum(scc$dr1.5)
scc2 <- sum(scc$dr2)
scc3 <- sum(scc$dr3)
scc4 <- sum(scc$dr4)
scc5 <- sum(scc$dr5)
scc6 <- sum(scc$dr6)
scc_vector <- c(scc0.025, scc0.5, scc0.75, scc1, scc1.5, scc2, scc3, scc4, scc5, scc6)
dr <- c(0.25, 0.5, 0.75, 1, 1.5, 2, 3, 4, 5, 6)
scc_table <- data.frame(scc_vector, dr)

ggplot(data = scc_table, aes(x = dr, y = scc_vector)) +
  geom_line() +
  geom_col(size = 3, fill = "firebrick2") +
  labs(x = "Discount rate (%)", y = "Social cost of carbon") +
  theme_minimal(14)
```


## 4. The National Academies of Sciences, Engineering, and Medicine advised the government in a 2017 report to use the Ramsey Rule when discounting within the SCC calculation: r = ρ + ηg Using ρ = 0.001, η = 2, and g = 0.01, what is the SCC? Locate this point on your graph from above.

```{r}
# find the new r
#0.001 + 2(0.01) = 0.021


```

## 5. Now suppose there are two possible climate policies that can be pursued. Policy A is business as usual and Policy B is to take immediate and strong action on climate change. Use these facts
- If you undertake Policy A there are two possible outcomes. Either warming will occur as
in the “baseline” (i.e. “no-pulse”) dataset above (this happens with probability 0.5) or
warming each year will be 1.5 times that in the “baseline” dataset (with probability 0.5).

- Under Policy B, warming will continue until 2050 as in the “baseline” dataset, and then
will stabilize at 1.29 degrees and stay that way forever.

- Society is risk neutral

- Use a discount rate of 2%

## What is the expected present value of damages up to 2100 under Policy A? What is the expected present value of damages up to 2100 under Policy B? Suppose undertaking Policy A costs zero and undertaking Policy B costs X. The total cost of a policy is the implementation cost plus expected damages. Make a plot of total cost of Policy A and total cost of Policy B against X. Suppose Congress wants to undertake the policy with the lower total cost. Use your graph to advise Congress.

## 6. Now suppose that society is risk averse. How do you think this would change your advice to Congress?
